Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-25 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of overlapping intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As in
Table~\ref{tab:I_n}, to select the correct interval from these overlapping
intervals~$[\ell,r]$, we define the non-overlapping interior part
of a interval, $\textrm{Int}\left([\ell,r]\right)$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$, their interior, and
their length, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.
See Table~\ref{tab:I_2} for a concrete example when $n=2$.}

\begin{tabular}{cccc}
\toprule 
\addlinespace
$k$ & $[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$ & $r-\ell$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & $\imposs$ & $\imposs$ & $0$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0$ & $\left[0,\tfrac{1}{2n}\right]$ & $\left(0,\tfrac{0.5}{2n}\right]$ & $\tfrac{1}{2n}$\tabularnewline\addlinespace
\midrule 
\addlinespace
\begin{tabular}{c}
$0<k<2n$\tabularnewline
and $2\nmid k$\tabularnewline
\end{tabular} & $\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]$ & $\left(\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right)$ & $\tfrac{1}{n}$\tabularnewline\addlinespace
\midrule 
\addlinespace
\begin{tabular}{c}
$0<k<2n$\tabularnewline
and $2\mid k$\tabularnewline
\end{tabular} & \multirow{1}{*}{$\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]$} & \multirow{1}{*}{$\left[\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right]$} & $\tfrac{1}{n}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$2n$ & $\left[\tfrac{2n-1}{2n},1\right]$ & $\left[\tfrac{2n-0.5}{2n},1\right)$ & $\tfrac{1}{2n}$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & $\necess$ & $\necess$ & $0$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
 Table~\ref{tab:I_2} lists a special case $\mathscr{I}_{2}$ to
illustrate these intervals and their interior parts more clearly.
\begin{table}
\caption{\label{tab:I_2}Intervals in $\mathscr{I}_{2}$, their interior, and
their length. Although intervals~$[\ell,r]$ are overlapping, their
interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.}

\begin{tabular}{cccc}
\toprule 
\addlinespace
$k$ & $[\ell,r]$ & $\textrm{Int}\left([\ell,r]\right)$ & $r-\ell$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & $\imposs$ & $\imposs$ & $0$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0$ & $\left[0,\tfrac{1}{4}\right]$ & $\left(0,\tfrac{1}{8}\right]$ & $\tfrac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$1$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$ & $\left(\tfrac{1}{8},\tfrac{3}{8}\right)$ & $\tfrac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$2$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$ & $\left[\tfrac{3}{8},\tfrac{5}{8}\right]$ & $\tfrac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$3$ & $\left[\tfrac{1}{2},1\right]$ & $\left(\tfrac{5}{8},\tfrac{7}{8}\right)$ & $\tfrac{1}{2}$\tabularnewline\addlinespace
\midrule 
\addlinespace
$4$ & $\left[\tfrac{1}{4},1\right]$ & $\left[\tfrac{7}{8},1\right)$ & $\tfrac{1}{4}$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & $\necess$ & $\necess$ & $0$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
 Then, given a real-valued probability~$x$, we will select an interval~$\iota_{n}\left(x\right)=[\ell,r]$
if $x\in\textrm{Int}\left([\ell,r]\right)$.

@@

\begin{thm}[Finite-precision Extension of Gleason's theorem for the
Born QIVPMs]In a Hilbert space $\Hilb$ of dimension $3$, given
any totally mixed state~$\rho$, 
\begin{equation}
B_{\frac{1}{12n}}\left(\rho\right)\subseteq\coreBorn\left(\iota_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)\subseteq B_{1/n}\left(\rho\right)\,.
\end{equation}
\end{thm}

\noindent In another word, if a state~$\rho'$ is consistent with
$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$, its distance with $\rho$,
$\left\Vert \rho-\rho'\right\Vert $, must be bounded by $1/n$ which
is the maximal length of the intervals in $\mathscr{I}_{n}$~\citep{HOSTunpublished}.
Conversely, if the distance between $\rho$ and $\rho'$ is bounded
by $\frac{1}{12n}$, their Born quantum probability on any projector~$P$
is $\left|\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)-\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right|\le\frac{1}{4n}$~\citep{HOSTunpublished}.
Moreover, the interiors of the intervals in $\mathscr{I}_{n}$ shrinks
$\frac{0.5}{2n}$ from the both ends, excepts $\imposs$, $\left[0,\tfrac{1}{2n}\right]$,
$\left[\tfrac{2n-1}{2n},1\right]$, and $\necess$. Together with
the fact that $0<\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)<1$ and
$0<\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)<1$, $\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)$
and $\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)$ are both in the same
intervals, and $\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$.
For example, we should have $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the state very close to
$\rho$ are guaranteed to belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.

@@

\noindent Given a three-dimensional Hilbert space $\Hilb$ with an
orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let
$\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and \begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations} It is straightforward to verify that $\bar{\mu}$
is a QIVPM. Next we will prove $\coreBorn\left(\bar{\mu},\events\right)$
is an empty set by contradiction. Suppose there is a state $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any~$P$, the state~$\rho$ must satisfy: 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0\,,
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. The last equation never holds proving the theorem. Thus,
there is no state consistent with the QIVPM on all projectors~$P\in\events$,
so Gleason's theorem cannot be extended to general QIVPMs.

Despite the non-extensibility of Gleason's theorem to general QIVPMs,
when we restrict $\bar{\mu}$ on any commuting subspace of events~$\events_{C}$,
$\bar{\mu}$ becomes classical, and Thm.~\ref{thm:Shapley} guarantees
$\coreBorn\left(\bar{\mu},\events_{C}\right)$ is nonempty. In this
sense, Thm.~\ref{thm:Non-extensible-of-Gleason's} is a completely
new phenomenon when we generalize both classical IVPMs and quantum
probability measures to QIVPMs.

>>>>

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of overlapping intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As in
Table~\ref{tab:I_n}, to select the correct interval from these overlapping
intervals~$[\ell,r]$, we define the non-overlapping interior part
of a interval, $\textrm{Int}\left([\ell,r]\right)$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$, their lengths, and
their interiors, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.
See Table~\ref{tab:I_2} for a concrete example when $n=2$.}

\begin{tabular}{cccc}
\toprule 
\addlinespace
$k$ & $[\ell,r]$ & $r-\ell$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & $\imposs$ & $0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0$ & $\left[0,\tfrac{1}{2n}\right]$ & $\tfrac{1}{2n}$ & $\left(0,\tfrac{0.5}{2n}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
\begin{tabular}{c}
$0<k<2n$\tabularnewline
and $2\nmid k$\tabularnewline
\end{tabular} & $\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]$ & $\tfrac{1}{n}$ & $\left(\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
\begin{tabular}{c}
$0<k<2n$\tabularnewline
and $2\mid k$\tabularnewline
\end{tabular} & \multirow{1}{*}{$\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]$} & $\tfrac{1}{n}$ & $\left[\tfrac{k-0.5}{2n},\tfrac{k+0.5}{2n}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$2n$ & $\left[\tfrac{2n-1}{2n},1\right]$ & $\tfrac{1}{2n}$ & $\left[\tfrac{2n-0.5}{2n},1\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & $\necess$ & $0$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
 Table~\ref{tab:I_2} lists a special case $\mathscr{I}_{2}$ to
illustrate these intervals and their interior parts more clearly.
\begin{table}
\caption{\label{tab:I_2}Intervals in $\mathscr{I}_{2}$, , their lengths,
and their interiors. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.}

\begin{tabular}{cccc}
\toprule 
\addlinespace
$k$ & $[\ell,r]$ & $r-\ell$ & $\textrm{Int}\left([\ell,r]\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
 & $\imposs$ & $0$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$0$ & $\left[0,\tfrac{1}{4}\right]$ & $\tfrac{1}{4}$ & $\left(0,\tfrac{1}{8}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$1$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$ & $\tfrac{1}{2}$ & $\left(\tfrac{1}{8},\tfrac{3}{8}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
$2$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$ & $\tfrac{1}{2}$ & $\left[\tfrac{3}{8},\tfrac{5}{8}\right]$\tabularnewline\addlinespace
\midrule 
\addlinespace
$3$ & $\left[\tfrac{1}{2},1\right]$ & $\tfrac{1}{2}$ & $\left(\tfrac{5}{8},\tfrac{7}{8}\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
$4$ & $\left[\tfrac{1}{4},1\right]$ & $\tfrac{1}{4}$ & $\left[\tfrac{7}{8},1\right)$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & $\necess$ & $0$ & $\necess$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
 Then, given a real-valued probability~$x$, we will select the interval~$\iota_{n}\left(x\right)=[\ell,r]$
if $x\in\textrm{Int}\left([\ell,r]\right)$.

@@

\begin{thm}[Finite-precision Extension of Gleason's theorem for the
Born QIVPMs]\label{thm:Finite-precision-Gleason}In a Hilbert space
$\Hilb$ of dimension $3$, given any totally mixed state~$\rho$,
\begin{equation}
B_{\frac{1}{12n}}\left(\rho\right)\subseteq\coreBorn\left(\iota_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)\subseteq B_{1/n}\left(\rho\right)\,.
\end{equation}
\end{thm}

\noindent In another word, if a state~$\rho'$ is consistent with
$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$, the difference between the
Born quantum probabilities on any projector~$P$, $\left|\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)-\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right|$,
must be bounded by $1/n$ which is the maximal length of the intervals
in $\mathscr{I}_{n}$. So does the distance between $\rho$ and $\rho'$,
$\left\Vert \rho-\rho'\right\Vert $. Conversely, if the distance
between $\rho$ and $\rho'$ is bounded by $\frac{1}{12n}$, so does
the difference between their Born quantum probabilities on any 1-dimensioal
projector~$\proj{\psi}$, $\left|\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho_{1}}^{\mathrm{B}}\left(\proj{\psi}\right)\right|\le\frac{1}{12n}$.
Since the dimension of the Hilbert space is $3$, the same difference
on any projector~$P$, $\left|\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)-\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right|$,
is bounded by $\frac{1}{4n}$. Moreover, the interiors of the intervals
in $\mathscr{I}_{n}$ shrinks $\frac{0.5}{2n}$ from the both ends,
excepts $\imposs$, $\left[0,\tfrac{1}{2n}\right]$, $\left[\tfrac{2n-1}{2n},1\right]$,
and $\necess$. Together with the fact that $0<\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)<1$
and $0<\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)<1$, $\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)$
and $\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)$ are both in the same
intervals, and $\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$.

As a concrete example of Thm.~\ref{thm:Finite-precision-Gleason},
we should have $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the state very close to
$\rho$ are guaranteed to belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.

@@

\noindent Given an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,
let $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a non-Born QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space.}

\begin{tabular}{cc}
\toprule 
\addlinespace
$P$ & $\bar{\mu}\left(P\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$, $\proj{0}$, $\proj{\ps}$, $\proj{\ps'}$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$, $\mathbb{1}-\proj{0}$, $\mathbb{1}-\proj{\ps}$, $\mathbb{1}-\proj{\ps'}$ & $\necess$\tabularnewline\addlinespace
\midrule 
\addlinespace
All other projectors & $\unknown$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Next we will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an
empty set by contradiction. Suppose there is a state $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any~$P$, the state~$\rho$ must satisfy: 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0\,,
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. The last equation never holds proving the theorem. Thus,
there is no state consistent with the QIVPM on all projectors~$P\in\events$,
so Gleason's theorem cannot be extended to general QIVPMs.

Notice that if we restrict $\bar{\mu}$ on the commuting subspace
of events~$\events_{C}$, the state~$\frac{1}{2}\cdot\proj{1}+\frac{1}{2}\cdot\proj{2}\in\coreBorn\left(\bar{\mu},\events_{C}\right)$
so that $\coreBorn\left(\bar{\mu},\events_{C}\right)$ is non-empty.
In general, for any commuting subspace of events~$\events_{C}'$,
$\coreBorn\left(\bar{\mu},\events_{C}'\right)$ is non-empty according
to the classical Thm.~\ref{thm:Shapley}. The difference between
classical and quantum highlights the following fact: Although classically
we could build a theory and deal the finite-precision measurement
in error analysis, it is necessary to embed the finite-precision measurement
into the foundation of quantum theory.
